Applications of the Fresnel Integrals

### 2019-03-17 draft.tex
more words

\section{Introduction}
An Euler spiral is a curve whose curvature changes linearly with its arc length. Taking the example of peeling an orange with a kitchen knife, one can either cut the skin along meridians, or cut it along a spiral as the left-hand image in Figure~\ref{f:orangePeel} shows. The red lines marked on the orange refer to the route one can follow to peel the orange. Later, if we unfold the spiral strip and flatten it on a table, we will see a long curve which is curved differently at different positions. Imagine that we will now cut the peel with progressively thinner strip widths, what we will obtain in the end is a long beautiful curve known as Euler spiral, as the right-hand image in Figure~\ref{f:orangePeel} shows. According to Alfred Gray, Euler spiral is “one of the most elegant of all plane curves.” \cite{ASG17}

@@

For a mathematician, an interesting question is whether there are any equations that can describe the corresponding shape of an Euler spiral. For this we consider the Fresnel integrals.  These are defined as integrals over some elementary functions which can however not be integrated in elementary functions.  These integrals occur naturally in the description of the optics of an illuminated straight edge (see Subsection~\ref{s:relation}).

\subsection{Definition of Euler Spiral}
The Euler Spiral is defined by the  linear relationship between curvature and arc length and was
first proposed as a problem of elasticity by James Bernoulli, then solved accurately by Leonhard Euler \cite{Lev08}.
The Euler spiral is a curve in the complex plane that has curvature proportional to the arc length.  Here the arc length $s(t)$ of a differentiable curve $(x(t),y(t))$ in the plane is defined by
\[  \left(\udfrac{s}{t}\right)^2 := \left(\udfrac{x}{t}\right)^2 +\left(\udfrac{y}{t}\right)^2.
\]  Note that the arc length is monotone increasing with the parameter $t$ of the curve.  For a nice curve, it is possible to rewrite the curve in the arc length, i.e.~$(x(s),y(s))$.

@@

\subsection{Mathematical history}
The Fresnel integral and the related Euler spiral were discovered multiple times during history.  One of the first occurrences is according to \cite{Lev08} by J. Bernoulli in 1694 where he defines a curve such that an ideal metal rod of this shape unwinds to a straight line when placed under load at the ends.  In 1744, Euler is able to describe the curve by a differential equation that also allows him to reformulate it as the Fresnel integral.  37 years later he is able to compute the limiting points ($t\to\pm\infty$) of the curve and integrals.  Fresnel discovers the curve in 1818 independently and successfully applies it to the diffraction on a slit \cite{Lev08}.  He is also able to compute a couple of values numerically.  In 1874 Cornu plotted the curve accurately and proposes it as a graphical means for computing diffraction problems.

In the following sections, we will first describe the important properties of the Fresnel integrals followed by some important applications of it.

\section{Properties of the Fresnel Integrals}
\subsection{Basic properties}
The Fresnel integrals are odd, i.e.
\begin{align*}
	\S(-x) &= \S(x), & \C(-x) &= \C(x).
\end{align*}
\begin{proof}[Idea of proof]  The integrands $\sin(t^2)$ and $\cos(t^2)$ are even in $t$ and an anti-derivative $F(x)$ of an even function is odd if in addition $F(0)=0$.  The latter follows, because both integrals start at $t=0$.
\end{proof}

@@

\subsection{Fresnel Diffraction}
\defbf\E{E}
\defbf\B{B}
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\textwidth]{opticalSetting.png} \\
	% (2) \includegraphics[width=0.5\textwidth]{opticalParameters.png}
	\label{f:FresnelDiffraction}
	\caption{Fresnel diffraction after \cite[ch.~8.1]{Zim08}. Incoming parallel light waves from the left.  An aperture S' and an observation screen on the right in distance $D$.}
\end{figure}
According to \cite[ch.~7]{Zim08} the complexified electric field strength $\E$ at a point $P$ on the observation screen of a plane electromagnetic wave that went through an aperture $S'$ computes as
\[  \E(P) = \B\iint  \frac1\rho \exp\left(\si(k\rho-\omega t)\right) \d{S'}
\] where $\B$ determines the amplitude of the incoming wave,\\ $\rho=\sqrt{(x-x')^2+(y-y')^2+D^2}$ is the distance of the aperture point $S'$ from the observation point $P$.  $k=\omega/c$ is the wave number, $\omega>0$ the frequency of the incoming monochromatic wave and $c>0$ the speed of light.  In the Fresnel approximation we assume that $|x-x'|/D\ll 1$ and $|y-y'|/D\ll 1$, but will not assume that the aperture itself is small.  Therefore we expand in the lowest orders of $(x-x')/D$ and $(y-y')/D$ and obtain
\[  \E(P) \approx \B\frac{\exp\left(\si(kD-\omega t)\right)}{D} \iint \exp\left(\si(x'-x)^2k/2D +\si(y'-y)^2k/2D\right) \d{(x',y')}.
\]  If we combine the constant coefficient as $\E_0$ and assume a rectangular aperture of size $[x_0,x_1]\times[y_0,y_1]$, we obtain
\begin{align*}  \E(x,y) &\approx \E_0 \int_{u_0}^{u_1} \exp\left(\si\frac\pi2 u^2\right)\d{u} \int_{v_0}^{v_1} \exp\left(\si\frac\pi2 v^2\right) \d{v} \\
  &= \E_0 \left[\C(u_1)+\si\S(u_1)-\C(u_0)-\si\S(u_0)\right] \left[\C(v_1)+\si\S(v_1)-\C(v_0)-\si\S(v_0)\right]
\end{align*} where $u_k:=\sqrt{\frac2{D\lambda}}(x_k-x)$, $v_k:= \sqrt{\frac2{D\lambda}}(y_k-y)$ for $k=0,1$ \cite[Eqns.~(8.34a\&b)]{Zim08} and $\lambda=2\pi/k$ is the wave length.  The intensity near a vertical edge results as
\[  I(x):= |\E|^2 = \frac{I_0}{2} \left[\left(\frac12-\C(u_0)\right)^2+\left(\frac12-\S(u_0)\right)^2\right].
\] See also \cite[Eqn.~(8.38)]{Zim08}.  This formula is the distance along the Euler spiral between the right point at $t=\infty$ and the left point at $t=u_0$.

>>>>

\section{Introduction}
An Euler spiral is a famous curve whose curvature changes proportional to  its arc length. Taking the example of peeling an orange with a kitchen knife, one can either cut the skin vertically, or cut it along a spiral as the left-hand image in Figure~\ref{f:orangePeel} shows. The red lines marked on the orange refer to the route one can follow to peel the orange. Later, if we unfold the spiral strip and flatten it on a table, we will see a long curve which is curved differently at different positions. Imagine that we will now cut the peel with progressively thinner width of the strip, what we will obtain in the end is a long beautiful curve known as Euler spiral, as the right-hand image in Figure~\ref{f:orangePeel} shows. According to Alfred Gray, Euler spiral is “one of the most elegant of all plane curves.” \cite{ASG17}

@@

For a mathematician, an interesting question is whether there are any equations that can describe the corresponding shape of an Euler spiral. For this we consider the Fresnel integrals.  These are defined as integrals over some elementary functions which can however not be integrated in elementary functions.  These integrals occur naturally in the description of the optics of an illuminated straight edge (see Subsection~\ref{s:app1}).

\subsection{Definition of Euler Spiral}
The Euler Spiral, defined by the  linear relationship between curvature and arc length, was
first proposed as a problem of elasticity by James Bernoulli, then solved accurately by Leonhard Euler \cite{Lev08}.
The Euler spiral is a curve in the complex plane that has curvature proportional to the arc length.  Here the arc length $s(t)$ of a differentiable curve $(x(t),y(t))$ in the plane is defined by
\[  \left(\udfrac{s}{t}\right)^2 := \left(\udfrac{x}{t}\right)^2 +\left(\udfrac{y}{t}\right)^2.
\]  Note that the arc length is monotone increasing with the parameter $t$ of the curve.  For a nice curve, it is possible to rewrite the curve in the arc length, i.e.~$(x(s),y(s))$.

@@

\subsection{Mathematical history}
If we trace the history of the Euler spiral and Fresnel integrals, we will find it under different authors, e.g.~Fresnel, Cornu, Euler etc.  This is because of the long history that Fresnel integral and the related Euler spiral have already left behind. They were discovered multiple times and under varying names during history.  One of the first occurrences is according to \cite{Lev08} by the French physicist J. Bernoulli in 1694.  Bernoulli approached many problems in elasticity theory at that time. One of them is a curve such that an ideal metal rod of this shape unwinds to a straight line when placed under load at the ends. He tried to find an explicit mathematical description of that curve, but it turned out that it is not that easy to do so at that time. In 1744, Swiss physicist and mathematician L. Euler is able to describe the same curve by a differential equation. This allows him to formulate a solution with integrals, namely what will later be called the Fresnel integrals.  37 years later, he computes the limiting points ($t\to\pm\infty$) of the curve and integrals which are $(\pm1/2,\pm1/2)$. In 1818, the French A.J. Fresnel who deals with many problems in wave optics and is unaware of the earlier results by Bernoulli and Euler discovers the Euler spiral independently and successfully applies it to the diffraction on a slit \cite{Lev08}.  He is also able to compute a couple of values numerically. With this finding he is able to establish the success of wave optics over particle optics which describes light as a transversal wave as opposed to the older idea by Newton who described light as consisting of particles.  In 1874 the French physicist M.A. Cornu followed similar ideas as Fresnel and is able to plot the curve accurately and proposes it as a graphical means for computing diffraction problems.  For that reason the Euler spiral is often also called Cornu spiral.

\section{Properties of the Fresnel Integrals}
The Fresnel integrals have many interesting mathematical properties. In this section, we will present some of them. 
\subsection{Basic properties}
The Fresnel integrals are odd, i.e.
\begin{align*}
	\S(-x) &= \S(x), & \C(-x) &= \C(x).
\end{align*}
\begin{proof}[Idea of proof]  The integrands $\sin(t^2)$ and $\cos(t^2)$ are even in $t$ and an anti-derivative $F(x)$ of an even function is odd if in addition $F(0)=0$.  The latter follows, because both integrals start at $t=0$.
\end{proof}

@@

\subsection{Fresnel Diffraction}\label{s:app1}
\defbf\E{E}
\defbf\B{B}
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\textwidth]{opticalSetting.png} \\
	% (2) \includegraphics[width=0.5\textwidth]{opticalParameters.png}
	\label{f:FresnelDiffraction}
	\caption{Fresnel diffraction after \cite[ch.~8.1]{Zim08}. Incoming parallel light waves from the left.  An aperture S' and an observation screen on the right in distance $D$.}
\end{figure}
According to \cite[ch.~7]{Zim08} the complexified electric field strength $\E$ at a point $P$ on the observation screen of a plane electromagnetic wave that went through an aperture $S'$ computes as
\[  \E(P) = \B\iint  \frac1\rho \exp\left(\si(k\rho-\omega t)\right) \d{S'}
\] where $\B$ determines the amplitude of the incoming wave,\\ $\rho=\sqrt{(x-x')^2+(y-y')^2+D^2}$ is the distance of the aperture point $S'$ from the observation point $P$.  $k=\omega/c$ is the wave number, $\omega>0$ the frequency of the incoming monochromatic wave and $c>0$ the speed of light.  In the Fresnel approximation we assume that $|x-x'|/D\ll 1$ and $|y-y'|/D\ll 1$, but will not assume that the aperture itself is small.  Therefore we expand in the lowest orders of $(x-x')/D$ and $(y-y')/D$ and obtain
\[  \E(P) \approx \B\frac{\exp\left(\si(kD-\omega t)\right)}{D} \iint \exp\left(\si(x'-x)^2k/2D +\si(y'-y)^2k/2D\right) \d{(x',y')}.
\]  If we combine the constant coefficient as $\E_0$ and assume a rectangular aperture of size $[x_0,x_1]\times[y_0,y_1]$, we obtain
\begin{align*}  \E(x,y) &\approx \E_0 \int_{u_0}^{u_1} \exp\left(\si\frac\pi2 u^2\right)\d{u} \int_{v_0}^{v_1} \exp\left(\si\frac\pi2 v^2\right) \d{v} \\
  &= \E_0 \left[\C(u_1)+\si\S(u_1)-\C(u_0)-\si\S(u_0)\right] \left[\C(v_1)+\si\S(v_1)-\C(v_0)-\si\S(v_0)\right]
\end{align*} where $u_k:=\sqrt{\frac2{D\lambda}}(x_k-x)$, $v_k:= \sqrt{\frac2{D\lambda}}(y_k-y)$ for $k=0,1$ \cite[Eqns.~(8.34a\&b)]{Zim08} and $\lambda=2\pi/k$ is the wave length.  The intensity near a vertical edge results as
\[  I(x):= |\E|^2 = \frac{I_0}{2} \left[\left(\frac12-\C(u_0)\right)^2+\left(\frac12-\S(u_0)\right)^2\right].
\] See also \cite[Eqn.~(8.38)]{Zim08}.  This formula is the distance along the Euler spiral between the right point at $t=\infty$ and the left point at $t=u_0$.
